### 2021-10-11 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Fixed commas, fullstops etc. that follow equations to make consistent.

A review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results. In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}. The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
\bseqn{
	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
}
where the radial part is
\bseqn{
	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
}
for $n - m$ even, while it is defined as $0$ for $n - m$ odd. A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}. These radial OPs are:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$. These OPs constitute a one-parameter family of orthogonal polynomials on the unit disk, with the classical Zernike polynomials being retrieved via setting the parameter $k = 0$. The use of polar coordinates allows for radially symmetric partial differential operators, while the incorporation of a parameter means they can be represented as banded matrices by exploiting the relations of the Jacobi polynomials. Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
\bseqnnumber{
	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:ds:triangleops}
}
for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values. Since there is no radial symmetry, differential operators for the triangle (as will be the case for the domains we investigate in this chapter) will be more complex than those for the disk.

@@

More precisely, we will consider the solution of partial differential equations on the domain
\bseqn{
	\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)} 
}
where  either of the following conditions hold:

@@

We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}. On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
\bseqn{
	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c
}
for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$. The corresponding OPs are denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
\bseqn{
	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
}
where $C_{n,k} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.

@@

\begin{proposition}[{\cite[p.55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then, for $n = 0,1,2,\dots$, $k = 0,\dots,n$, the 2D polynomials defined on $\Omega$ given by
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)})
}
are orthogonal polynomials with respect to the weight $W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)})$ on $\Omega$. 
\end{proposition}

For disk slices and trapeziums, we specialise \bsrefprop{prop:ds:construction} in the following definitions. First we introduce notation for two families of univariate OPs. \nomenclature[weight]{$\genjacw^{(a,b,c)}$, $\jacw^{(a,b)}$}{Weight functions for the one-dimensional OPs $\{\genjac_n^{(a,b,c)}\}$ and $\{\jac_n^{(a,b)}\}$ respectively} 
\begin{definition}\label{def:ds:OPconstruction}
	Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
\bseqn{
	\begin{cases}
		\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
		\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b
	\end{cases}
}
and define the associated inner products by: \nomenclature[omega]{$\normgenjac$, $\normjac$}{Normalising constants for inner products} \nomenclature[innerw]{$\ip< p, \: q >_w$}{A one-dimensional inner product with weight $w$}
\bseqnnumber{
	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ds:ipgenjac} \\
	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ds:ipjac}
}
where
\bseqnnumber{
	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ds:ipnormalisations}
}
Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by \nomenclature[R]{$\genjac_n^{(a,b,c)}$}{Non-classical three-parameter univariate OP of degree $n$} $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjac}, and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipjac}.
\end{definition}

We can now write down our 2D OP family. \nomenclature[Hnk]{$\hdopnk^{(a,b,c,d)}$}{Four-parameter bivariate OP of degree $n$ and order $k$ defined on the given domain and by the given inner product}
\begin{definition}\label{def:ds:constuction}
	For parameters $a,b,c,d \in \R$, define the four-parameter 2D orthogonal polynomials as the set $\{\hdopnk^{(a,b,c,d)}\}$ where:
\bseqn{
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
}
\end{definition}

@@

$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight \nomenclature[Wa]{$W^{(a)}, W^{(a,b,c)}, W^{(a,b,c,d)}$}{Weight function for relevant multidimensional OPs, with the given parameters}
\bseqn{
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
}
assuming that either \bsrefcondition{cond:trap} holds, or \bsrefcondition{cond:disk} holds with $\jacw^{(a,b)}$ being an even function (i.e. $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) = (\delta-x^2)^a$). That is,
\bseqn{
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{\hdop^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}
}
where for $f, g : \Omega \to \R$ the inner product is defined as \nomenclature[innerH]{$\ip< p, \: q >_{\hdopabcd}$}{A multidimensional inner product over $\Omega$ with the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
\bseqn{
	\ip< f, \: g >_{\hdop^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
}

@@

For the disk-slice, the weight $\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c$ results from setting:
\bseqn{
\begin{cases}
	(\alpha,\beta) &\subset (0,1) \\
	(\gamma,\delta) &:= (-1,1) \\
	\rho(x) &:= (1-x^2)^{\half}
\end{cases}
}
so that
\bseqn{
\begin{cases}
	\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c} \\
	\jacw^{(c)}(x) := (1-x)^c \: (1+x)^c = (1-x^2)^c.
\end{cases}
}
Note here we can simply remove the need for including a fourth parameter $d$. The 2D OPs orthogonal with respect to the weight above on the disk-slice $\Omega$ are then given by:
\bseqnnumber{\label{eq:diskpolys}
	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
}
In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).

@@

Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method. To obtain their entries, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$. These can be expressed as: 
\bseqnnumber{
	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:ds:Hrecurrence} \\
	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
}
Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$. We can use \bsrefeqns{eqn:ds:Hrecurrence}{eqn:ds:Precurrence} to determine the 2D recurrences for $\hdopnkabcd(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships.

>>>>

A review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results. In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}. The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
\bseqn{
	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
}
where the radial part is
\bseqn{
	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1),
}
for $n - m$ even, while it is defined as $0$ for $n - m$ odd. A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}. These radial OPs are:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$. These OPs constitute a one-parameter family of orthogonal polynomials on the unit disk, with the classical Zernike polynomials being retrieved via setting the parameter $k = 0$. The use of polar coordinates allows for radially symmetric partial differential operators, while the incorporation of a parameter means they can be represented as banded matrices by exploiting the relations of the Jacobi polynomials. Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
\bseqnnumber{
	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}), \label{eqn:ds:triangleops}
}
for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values. Since there is no radial symmetry, differential operators for the triangle (as will be the case for the domains we investigate in this chapter) will be more complex than those for the disk.

@@

More precisely, we will consider the solution of partial differential equations on the domain
\bseqn{
	\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)},
}
where either of the following conditions hold:

@@

We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}. On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
\bseqn{
	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c,
}
for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$. The corresponding OPs are denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
\bseqn{
	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms}),
}
where $C_{n,k} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.

@@

\begin{proposition}[{\cite[p.55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then, for $n = 0,1,2,\dots$, $k = 0,\dots,n$, the 2D polynomials defined on $\Omega$ given by
\bseqn{
	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}),
}
are orthogonal polynomials with respect to the weight $W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)})$ on $\Omega$. 
\end{proposition}

For disk slices and trapeziums, we specialise \bsrefprop{prop:ds:construction} in the following definitions. First we introduce notation for two families of univariate OPs. \nomenclature[weight]{$\genjacw^{(a,b,c)}$, $\jacw^{(a,b)}$}{Weight functions for the one-dimensional OPs $\{\genjac_n^{(a,b,c)}\}$ and $\{\jac_n^{(a,b)}\}$ respectively} 
\begin{definition}\label{def:ds:OPconstruction}
	Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
\bseqn{
	\begin{cases}
		\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c}, \\
		\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b,
	\end{cases}
}
and define the associated inner products by: \nomenclature[omega]{$\normgenjac$, $\normjac$}{Normalising constants for inner products} \nomenclature[innerw]{$\ip< p, \: q >_w$}{A one-dimensional inner product with weight $w$}
\bseqnnumber{
	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x, \label{eqn:ds:ipgenjac} \\
	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y, \label{eqn:ds:ipjac}
}
where
\bseqnnumber{
	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ds:ipnormalisations}
}
Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by \nomenclature[R]{$\genjac_n^{(a,b,c)}$}{Non-classical three-parameter univariate OP of degree $n$} $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjac}, and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipjac}.
\end{definition}

We can now write down our 2D OP family. \nomenclature[Hnk]{$\hdopnk^{(a,b,c,d)}$}{Four-parameter bivariate OP of degree $n$ and order $k$ defined on the given domain and by the given inner product}
\begin{definition}\label{def:ds:constuction}
	For parameters $a,b,c,d \in \R$, define the four-parameter 2D orthogonal polynomials as the set $\{\hdopnk^{(a,b,c,d)}\}$ where:
\bseqn{
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega.
}
\end{definition}

@@

$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight \nomenclature[Wa]{$W^{(a)}, W^{(a,b,c)}, W^{(a,b,c,d)}$}{Weight function for relevant multidimensional OPs, with the given parameters}
\bseqn{
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
}
assuming that either \bsrefcondition{cond:trap} holds, or \bsrefcondition{cond:disk} holds with $\jacw^{(a,b)}$ being an even function (i.e. $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) = (\delta-x^2)^a$). That is,
\bseqn{
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{\hdop^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
}
where for $f, g : \Omega \to \R$ the inner product is defined as \nomenclature[innerH]{$\ip< p, \: q >_{\hdopabcd}$}{A multidimensional inner product over $\Omega$ with the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
\bseqn{
	\ip< f, \: g >_{\hdop^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
}

@@

For the disk-slice, the weight $\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c$ results from setting:
\bseqn{
	\begin{cases}
		(\alpha,\beta) &\subset (0,1) \\
		(\gamma,\delta) &:= (-1,1) \\
		\rho(x) &:= (1-x^2)^{\half}
	\end{cases}
}
so that
\bseqn{
	\begin{cases}
		\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c} \\
		\jacw^{(c)}(x) := (1-x)^c \: (1+x)^c = (1-x^2)^c.
	\end{cases}
}
Note here we can simply remove the need for including a fourth parameter $d$. The 2D OPs orthogonal with respect to the weight above on the disk-slice $\Omega$ are then given by:
\bseqnnumber{\label{eq:diskpolys}
	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega.
}
In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).

@@

Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method. To obtain their entries, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$. These can be expressed as: 
\bseqnnumber{
	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x), \label{eqn:ds:Hrecurrence} \\
	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
}
Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$. We can use \bsrefeqns{eqn:ds:Hrecurrence}{eqn:ds:Precurrence} to determine the 2D recurrences for $\hdopnkabcd(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships.

\begin{lemma}
	$\hdopnkabcd(x,y)$ satisfy the following relations:
\bseqn{
	x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
	y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
}
for $(x,y) \in \Omega$, where
\bseqn{
	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)}, \\
	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}, \\
	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}, \\
	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}.
}
